---
title: "Dynamic Linear Model and Ideal Point Estimation"
author: "Guilherme Jardim Duarte"
date: "17 de junho de 2016"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is a case study and tutorial regarding how to write an Ideal Points Model which varies in time for Political Science in Political Science. It consists of four parts. Firsly, we will explain what is an Ideal Point Model. We will replicate the paper "Dynamic Ideal Point Estimation via Markov Chain Monte Carlo for the U.S. Supreme Court, 1953–1999", by Andrew D. Martin and Kevin M. Quinn, published in Political Analysis - 2001.

# Introduction 

## Ideal Point Modeling

Ideal Point Models are a common model for Political Science. They assume that politicans have preferences represented by utility functions to a vectorial space of $\mathbb{R}$. So, for example, we have two congressmen A and B, and they have to decide about one issue, we could think about this situation this way:

```{r, echo=FALSE}
library(ggplot2)
```

